### 2012-10-12 formalism/paper.tex
added updated version of the FOOL paper including discussion of the new type system amendments. Implementation includes join constraints but stubs them for the solver.

\documentclass[preprint]{sigplanconf}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}
\usepackage{listings}
\usepackage{proof}
\usepackage{supertabular}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{theorem}
\usepackage{url}
\usepackage{graphicx}
\usepackage{mathtools}

\newtheorem{lem}{Lemma}
\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}

\newcommand{\Tr}{\mathit{Tr}}
\newcommand{\secref}[1]{Section \ref{#1}}
\newcommand{\figref}[1]{Figure \ref{#1}}
\newcommand{\lstref}[1]{Listing \ref{#1}}
\newcommand{\appref}[1]{Appendix \ref{#1}}
\newcommand{\lemref}[1]{Lemma \ref{#1}}
\newcommand{\corref}[1]{Corollary \ref{#1}}

\newcommand{\rlett}{\textsc{r\_let\_term} }
\newcommand{\rletv}{\textsc{r\_let\_value} }
\newcommand{\rseqlt}{\textsc{r\_seq\_left\_term} }
\newcommand{\rseqlv}{\textsc{r\_seq\_left\_value} }
\newcommand{\rfunc}{\textsc{r\_fun\_call} }
\newcommand{\rmethc}{\textsc{r\_meth\_call} }
\newcommand{\rifexpr}{\textsc{r\_if\_term} }
\newcommand{\riftrue}{\textsc{r\_if\_true} }
\newcommand{\riffalse}{\textsc{r\_if\_false} }

\newcommand{\tunit}{\textsc{t\_unit} }
\newcommand{\ttrue}{\textsc{t\_true} }
\newcommand{\tfalse}{\textsc{t\_false} }
\newcommand{\tobj}{\textsc{t\_object} }
\newcommand{\tfundef}{\textsc{t\_fun\_def} }
\newcommand{\tlet}{\textsc{t\_let} }
\newcommand{\tif}{\textsc{t\_if} }
\newcommand{\tseq}{\textsc{t\_seq} }
\newcommand{\tfunc}{\textsc{t\_fun\_call} }
\newcommand{\tmethc}{\textsc{t\_meth\_call} }
\newcommand{\tdrop}{\textsc{t\_drop} }
\newcommand{\tsub}{\textsc{t\_sub} }
\newcommand{\tnarrow}{\textsc{t\_narrow} }
\newcommand{\twiden}{\textsc{t\_widen} }

\newcommand{\subrefl}{\textsc{sub\_refl}}
\newcommand{\subtrans}{\textsc{sub\_trans}}
\newcommand{\subfn}{\textsc{sub\_fn}}
\newcommand{\subobj}{\textsc{sub\_obj}}

\newcommand{\remapunit}{\textsc{remap\_unit}}
\newcommand{\remapbool}{\textsc{remap\_bool}}
\newcommand{\remapfunc}{\textsc{remap\_func}}
\newcommand{\remapobj}{\textsc{remap\_obj}}

\newcommand{\remapapp}[3]{\ensuremath{\mathbf{remap}(#1, #2 \gg #3)}}

\newcommand{\typerule}[4]{#1 \triangleright #2 : #3 \triangleleft #4}
\newcommand{\oprule}[4]{#1 \mid #2\;\longrightarrow\;#3 \mid #4}
\newcommand{\inferrule}[4]{\left\llbracket #1 \triangleright #2 : #3 \triangleleft #4 \right\rrbracket}
\newcommand{\inferlhs}[1]{\left\llbracket \gamma \triangleright #1 : \alpha \triangleleft \gamma' \right\rrbracket}
\newcommand{\subst}[3]{#3 \{\overline{^{#1}/_{#2}}\}}
\newcommand{\substxy}[1]{\subst{x_i}{y_i}{#1}}
\newcommand{\unitv}{\mathtt{unit}}
\newcommand{\unitt}{\mathbf{Unit}}
\newcommand{\truev}{\mathtt{true}}
\newcommand{\falsev}{\mathtt{false}}
\newcommand{\boolt}{\mathbf{Bool}}
\newcommand{\ot}[2]{#1@\overline{#2}}
\newcommand{\funv}[4]{\lambda(\overline{#1 : #2 \gg #3}).#4}
\newcommand{\funt}[3]{(\overline{#1 \gg #2}) \rightarrow #3}
\newcommand{\lett}[3]{\mathtt{let}\:#1\:\mathtt{=}\:#2\:\mathtt{in}\:#3}
\newcommand{\ift}[3]{\mathtt{if} \; #1 \; \mathtt{then} \; #2 \; \mathtt{else} \; #3}
\newcommand{\cand}{\:\wedge\:}
\newcommand{\free}[1]{\mathbf{free}\:#1}
\newcommand{\fresh}[1]{\mathbf{fresh}\:#1}

\newcommand{\qed}{$\blacksquare$}

\newenvironment{proof}{\vspace{1ex}\noindent{\bf Proof}\hspace{0.5em}}
  {\hfill\qed\vspace{1ex}}

\lstset{
    basicstyle=\ttfamily \small,
    numbers=left,
    frame=single,
    tabsize=2,
    xleftmargin=2.0em,
    xrightmargin=0.4em
}

\input{ts_defns}

\begin{document}

\conferenceinfo{TBD}{TBD} 
\copyrightyear{2012} 
\copyrightdata{[to be supplied]}

\titlebanner{IN-PROGRESS WORK}        % These are ignored unless
\preprintfooter{DO NOT REDISTRIBUTE WITH PERMISSION}   % 'preprint' option specified.

\title{Typestate Inference in an Imperative Calculus}

\authorinfo{Iain McGinniss \and Simon Gay}
           {University of Glasgow}
           {\{iainmcgin, simon\}@dcs.gla.ac.uk}

\maketitle

\begin{abstract}
Current typestate research is primarily focused on studying
type systems that can support particular use cases for typestate;
the type systems are often complex, and programs in such languages 
require large type annotations to allow for full static type checking. Little
work has been done to determine whether inference of such annotations is 
feasible. We present a small imperative calculus which
models interactions with typestate-constrained objects and a type inference
technique which can derive the minimal state machine that governs interactions
with the observed objects, and therefore eliminate the need for
explicit type annotations.
\end{abstract}

\category{D.2.4}{Software Engineering}{Programming by contract}
\category{F.3.3}{Studies of Program Constructs}{Type structure}
\category{D.1.5}{Object-oriented Programming}{}
\category{D.3.1}{Programming Languages}{Formal Definitions and Theory}

\keywords
typestate, type inference, constraint solving, alias control

\section{Introduction}

Typestate is ubiquitous in object oriented programs \cite{Beckman2011} and yet
very few languages provide direct support for it. Stateful objects are a source
of subtle runtime errors when their contracts are not respected, as the
point of failure is often separated from the cause in both source location
and time. As such the use of immutable objects and the 
borrowing of ideas from functional programming has become much more prevalent, 
such as in languages like Scala and Rust \cite{Rust}.

Immutability runs contrary to what objects were intended to be --- objects in
their most fundamental form can be viewed as entities with private state and 
associated protocols, which specify the messages the entities can respond to.
An immutable object, meaning an object with constant private state and
a stateless protocol, has much more in common with a type constructor in 
a functional language.

An object may have a very simple 
protocol, where the set of messages (or equivalently, method calls) it is 
willing to accept is constant throughout its lifecycle --- this is what the 
type system of programming languages such as Java can express. This is
insufficient to correctly model the object's behaviour in many cases, as
the set of messages the object can correctly respond to is dependent upon an
abstraction of its private state. State change is typically synchronous, in 
response to message receipt, and safety requires that 
any message sent to an object is within the set which can be received for
the currently known state.

Describing the allowed sequence of messages which a class of objects can
receive, their typestate, produces significantly more complex types than
in an object oriented language without typestate constraints. Functions which
take objects as parameters must describe the type of object they accept,
in which state, and what state transformation will occur (which may not
be deterministic). This type annotation burden is particularly pronounced
in a language with function literals, which must be succinct
in order to be readable.

Additionally, static checking of typestate requires
that the current state of an object always be known at a point of invocation.
This is only possible when an alias control strategy is employed, such as linear types, 
but such types prohibit programming in a style that most programmers are 
familiar with. Attempting to relax linear typing through the use of fractional 
permissions as in Plural and Plaid \cite{Aldrich2009,Bierhoff2009}
adds a significant amount of complexity to the type system. This results in 
large type annotations within programs to express both typestate and aliasing 
constraints when modular static type checking is desired.
If it were possible to eliminate the need for such type annotations, code
could be more concise and type annotations would only be necessary for
documentation purposes, and could be checked that they are simply more specific
versions of the principal typing produced by the type inference algorithm.

\section{Modelling typestate constraints}
\label{sec:tsmodel}

In previous work \cite{McGinniss2011}, we have taken the view that most use 
cases for typestate can be adequately modelled by a deterministic
finite state machine. The state machine explicitly describes the
states of the object, the methods which can be called in each state and
the transition triggered by a method call in a specific state.
Fields on the object can either be treated as private or as a special case
of method invocation \cite{Abadi1996}. For simplicity
we will only consider methods with no parameters, as there are very few
use cases where the values of parameters have an effect on state transition 
decisions which could be described without the integration of much more
complex features into the model, such as range types, conditional transitions
or a full logical framework.

The type of a given object at any point in time can be described as the
combination of a state machine (which we will refer to as the 
{\it object protocol}) and a particular state from within that protocol.
We write this as $O@S$, where $O$ is the protocol and $S$ is the state.
Formally, an object protocol $O$ is defined as a triple
$\langle M, \Sigma, \Delta \rangle$ where $M$ is a set of all
method names, $\Sigma$ is the set of state labels and
$\Delta \subseteq \Sigma \times M \times \Sigma$ is the state
transition relation. Transitions for a given state must be unique, such that 
$(S, m, S') \in \Delta \implies \nexists S'' \:\mathrm{st.}\: S' \neq S'' 
\wedge (S, m, S'') \in \Delta$.

The subtyping relation $<:$ is defined in terms of a
{\it simulation preorder}. Given $O = \langle M, \Sigma, \Delta \rangle$ and
$O_2 = \langle M_2, \Sigma_2, \Delta_2 \rangle$, the simulation preorder
$\sim$
is defined to be the largest relation such that if $O@S \sim O_2@S_2$ then for all
$m$ and $S_4$ such that $(S_2, m, S_4) \in \Delta_2$ there exists an $S_3$ such that
$(S, m, S_3) \in \Delta$ and that $O@S_3 \sim O_2@S_4$.
Subtyping is defined such that $O@S <: O_2@S_2 \iff O@S \sim O_2@S_2$.

A perhaps more intuitive basis for subtyping
uses the notion of {\it trace inclusion}. The set of interaction traces 
$\Tr(O@S)$ of a type is the smallest set such that:

\[
\begin{tabular}{cc}
\infer{\epsilon \in \Tr(O@S)}{}
&
\infer{m . t \in \Tr(O@S)}
{(S, m, S') \in \Delta & t \in \Tr(O@S')}
\end{tabular}
\]

where $\epsilon$ is the empty trace, and $m . t$ is the 
prefixing of an existing trace with a method call.
Subtyping is therefore defined such that
$O@S <: O_2@S_2 \iff \Tr(O_2@S_2) \subseteq \Tr(O@S)$.

It is easy to
show that $O@S \sim O_2@S_2 \iff \Tr(O_2@S_2) \subseteq \Tr(O@S)$, so these
interpretations of subtyping can be used interchangeably wherever necessary.

Either view of subtyping is also easy to relate to the Liskov substitution 
principle \cite{Liskov1987} --- a type will accept every sequence of
method calls that a supertype would, therefore it is a {\it behavioural}
subtype.

\section{TS - an imperative calculus with typestate}

TS is a small imperative calculus which models interactions with 
typestate-constrained objects, supports function literals without 
implicit capture, and employs structural subtyping for function and object
types.
The grammar of the language is shown 
in \figref{fig:grammar}, with operational semantics in \figref{fig:opsemantics} 
and typing rules in \figref{fig:typerules}.
The language is designed primarily
to study typestate inference in imperative code, rather than to 
be practical, as it uses syntactic control of interference \cite{Reynolds1978} 
to avoid the issue of alias control entirely - there is no means within the
language to produce an alias to an object.

The language has the following values:

\begin{itemize}
\item {\tt unit}, the sole value of the {\tt Unit} type, which is used as
the result type for terms which have no meaningful result.
\item Function literals, which do not allow for implicit capture
of variables into their scopes --- everything must be passed as a parameter.
Tracking effects on implicitly captured references is not very well
studied in the current literature outside of strict linear types where
function literals that capture linear references must themselves be linear.
Parameters are passed by reference. 
Requiring that all references be explicitly passed to the function allows for
effects to be explicitly declared. The notation $T \gg T'$ is used to
describe an {\it effect type}, which declares a required type for the parameter 
on the left, and the new type of the parameter after application of the function 
on the right. Not all effect types are legal, as will be discussed later.
\item Object literals. These are composed of a state machine description
and a current state. The state machine description contains a set of
distinct state labels $\overline{S_i}$, within which is a set of methods $\overline{m_{ij}}$ 
which can be called in that state. The language does
not attempt to fully model the behaviour of objects, only interactions with
them. As such, it is sufficient for each method to be defined as a pair 
consisting of the return value of the method, and the next state of the object 
if the method is invoked. Non-determinism in the production of a return value 
could be simulated by specifying a set of possible return values and allowing 
the operational semantics to select one of these randomly.
All values would have to have the same type, or types for which a least upper 
bound exists. Such non-determinism would have no effect on the type system,
which is the main focus of this paper.
\end{itemize}

\noindent
The language has the following terms:

\begin{itemize}
\item A conditional of form $\ift{t_c}{t_t}{t_f}$, which evaluates
$t_c$ first to a boolean result which decides which of $t_t$ 
(the \"true\" branch) or $t_f$ (the \"false\" branch) is
evaluated.
\item Function calls of form $x(\overline{x_i})$, which extract the function
literal stored in $x$ and apply it to the set of parameters $\overline{x_i}$,
which must match the declared parameter types. Each $x_i$ must be a distinct
variable, and is passed by reference, allowing the type of $x_i$ to be changed
as a result of the function application.
\item Method invocations on objects of form $x.m$. A method call is only
valid when the method is available in the current state of the object stored
in $x$. After a valid invocation, the state of the object will have changed
as specified by the object's state machine.
\item A let-bind construct $\lett{x}{t}{t'}$, which creates a new variable
$x$ within the scope of the body term $t'$ which is a reference to the value
derived from the term $t$. Rebinding a variable name is not allowed.
\item Sequencing of form $\tseq{t}{t'}$, which evaluates terms in 
left-to-right order.
\end{itemize}

While variables are references to a value, there is no construct which
can duplicate a reference. This very simple mechanism
ensures that references are unique, allowing us to focus our attention on the
typestate properties of objects without the added complication of alias
control annotations in the types. The restriction that each variable passed to
a function be unique is for this reason, to prevent implicit duplication
by passing the same variable more than once to the same function.

\begin{figure}

\ottgrammartabular{
\ottt\ottinterrule
\ottv\ottinterrule
\otto\ottinterrule
\ottT\ottinterrule
\ottO\ottinterrule
}

\caption{\label{fig:grammar} Grammar of TS0}
\end{figure}

The example in \figref{fig:exampleprog} shows a short TS0 program
that TODO.

\begin{figure}
\begin{lstlisting}
TODO
\end{lstlisting}
\caption{\label{fig:exampleprog} An example program in TS0}
\end{figure}

\subsection{Stores}

Each variable is a unique reference to a value in a store $\mu$, which is 
defined as $\mu\:\mathrm{::=}\:\overline{x_i \mapsto v_i}$ where each $x_i$ is
distinct. A store location
is created (or overwritten) upon entry to the body of a let-binding, written as
$\mu[x \mapsto v]$, which is defined as follows:

\[
\begin{tabular}{cc}
\infer
{\mu[x \mapsto v] = \mu, x \mapsto v}
{\mu = \overline{x_i \mapsto v_i} & x \notin \overline{x_i}}
&
\infer
{\mu[x \mapsto v] = \mu, x \mapsto v}
{\mu = \overline{x_i \mapsto v_i}, x \mapsto v'}
\end{tabular}
\]

As there is no way for a reference to be duplicated, the mapping
from variable names to store locations is injective and therefore we have
no need for explicit ``location values'' as in the typical treatment of 
languages with references (such as in \cite{Pierce2002}). There is no need
for a separate store typing, as a well-typed store can be defined directly
in relation to a context as in \figref{fig:storetype}. A well-typed store
may mappings for variables which are not in the context as shown in
rule \textsc{st\_extra}.

\begin{figure}

\ottdefnreduce

\caption{\label{fig:opsemantics} Operational semantics}
\end{figure}

\subsection{Type rules and soundness}

As the language is imperative and the types of variables pointing to objects can
change, it is necessary for the typing judgements to express changes to the 
environment. We must then be able
to prove that the changes in the environment correspond to the changes in the
store during evaluation. A context $\Gamma$ is a set of unique variable names 
mapped to types, defined as $\Gamma\:\mathrm{::=}\:\overline{x_i : T_i}$,
with each $x_i$ distinct. \figref{fig:storetype} defines what it means for
a store to be well-typed with respect to a
context. \figref{fig:typerules} defines the rules for well-typed terms
in the language, while \figref{fig:subtyperules} defines the subtyping
relation.

\begin{figure}
\begin{ottdefnblock}
{$\Gamma  \vdash  \mu$}
{\ottcom{store $\mu$ is well typed against context $\Gamma$}}
\end{ottdefnblock}

\[
\begin{tabular}{cc}
\infer[\textsc{st\_empty}]
{\emptyset \vdash \emptyset}
{}
&
\infer[\textsc{st\_vars}]
{\overline{x_i : T_i} \vdash \overline{x_i \mapsto v_i}}
{\typerule{\Gamma}{v_i}{T_i}{\Gamma'}}
\\\\
\multicolumn{2}{c}{
\infer[\textsc{st\_extra}]
{\Gamma \vdash \mu, \overline{y_j \mapsto v_j}}
{\Gamma \vdash \mu & \overline{y_j \notin dom(\mu)}}
}
\end{tabular}
\]

\caption{\label{fig:storetype} Store typing judgements}
\end{figure}

\begin{figure}
\ottdefnmethtype

\ottdefntype

\caption{\label{fig:typerules} Type rules}
\end{figure}

\begin{figure}

\ottdefnsubtype

\caption{\label{fig:subtyperules} Subtyping rules}
\end{figure}

Corresponding to the values of the language, there are four main kinds
in the type system: Unit, Bool, function types and object types. 
Two modifications are made to the definition of object protocols described
in \secref{sec:tsmodel}: methods have return types, and an object type
is the combination of an object protocol and a {\it set} of states.
An object protocol $O$ is therefore formally defined as the tuple 
$O = \langle M, \Sigma, \Delta \rangle$
where the transition relation $\Delta \subseteq \Sigma \times M \times T \times \Sigma$, 
where $T$ is the set of all types in the language.

While an object value in TS is still deterministic and has a single state at any 
given
point in time, we may not be able to determine in the type system which state 
exactly it is in --- conditional evaluation readily produces this situation.
With an if-then-else statement where the branches
interact with an object differently, it is common for the state of the object
at the end of each branch to be different. As
knowing precisely which branch will be taken is rarely possible in a
decidable type system, we must either insist that the state of the object
be the same upon exit from both branches, or devise a means to cope with
the object being in one of multiple potential states. We choose the latter
approach, as it is both more flexible and allows for intuitive principal
types to be chosen, as will be discussed later.

% abandoned attempt to define join/meet
% \begin{figure}

% \[
% \begin{array}{c}
% \infer[\sqcap-state]
%   {S \sqcup S_2 = \{ \overline{ m : T'' \Rightarrow (S_3 \sqcap S_4)} \} }
%   {
%   m : T \Rightarrow S_3 \in O@S
%   &
%   m : T' \Rightarrow S_4 \in O_2@S_2
%   &
%   T'' = T \sqcup T'
%   }
% \\
% \infer[\sqcup-omg]
%   {S \sqcup S_2 = 
%   \{ \overline{ m : T'' \Rightarrow (S_3 \sqcup S_4} \} } 
%   \cup
%   \{ \overline{ m' : T_3 \Rightarrow S_4 \} }
%   {}
% \]

% \caption{\label{fig:joinmeet} Join/meet definition}
% \end{figure}

An object type in TS, written $O@\overline{S}$, is equivalent to
$\bigsqcup_{S \in \overline{S}} O@S$, where $T \sqcup T'$ is the least upper
bound (or {\it join}) of $T$ and $T'$. Deriving a least upper bound for
two arbitrary object types (with a single current state) 
involves constructing the intersection of the state machines for each object 
type, using the current state of each object type as the initial state.
Similarly, deriving a greatest lower bound is based upon the union of
the two state machines in the same fashion. Join is always defined for
object types, while meet is not --- where the return types of methods are
incompatible (i.e. no join/meet exists respectively), a join operation will 
discard the method (it is unsafe to
call it) while a meet operation would require that the method be callable,
however no valid return type can be assigned to the method. Therefore
the kind of object types is a join semi-lattice, with the object type 
$\{ S_1 \{\} \}@S_1$ serving as the least element.

The intuition here is that the only sequences of method calls which can be 
safely invoked on an object of type $O@\overline{S}$ are those which could be 
safely invoked on each individual $O@S$, where $S \in \overline{S}$, and that
a least upper bound of the return types exists.

In general, the set of
types in TS is a disjoint union of join semi-lattices.

\subsection{Extra notational conventions}

Given the formal definition of an object type $O@\overline{S}$, to save
referring to the $\Sigma$, $M$ and $\Delta$ that $O$ is composed of we shall
use the following notational conventions:

\begin{itemize}
\item $S \in O$, a predicate asserting $S$ is a state of $O$, is equivalent to
$S \in \Sigma$.
\item $m : T \Rightarrow S' \in O@S$, the method existence predicate, 
states the method $m$ with return type $T$ exists in state $S$ and triggers a 
transition to state $S'$ when called. Formally, $(S,m,T,S') \in \Delta$.
\item $m : T \Rightarrow \overline{S'} \in O@\overline{S}$ is the extension
of the previous to state sets, meaning that for all $S \in \overline{S}$
there exists a state $S'$ and type $T'$ where 
$(S,m,T',S') \in \Delta$ and that
$T = \bigsqcup_{S \in \overline{S}} \{ T' | (S,m,T',S') \in \Delta \}$ exists
and that $\overline{S'} = \{ S' | (S,m,T',S') \in \Delta \}$.
\item $O@\overline{S} \sim O_2@\overline{S_2}$, the extension of the
simulation relation to support state sets, can be interpreted as
as $\bigsqcup_{S \in \overline{S}}O@S \sim \bigsqcup_{S_2 \in \overline{S_2}}O_2@S_2$.
\end{itemize}

\subsection{Effect types}

An ``effect type'' $T \gg T'$ describes how the type of a variable changes
in response to function application. In the TS language, the only type
changes that are those which change the state of an object instance; the
values of the other types are immutable. If a boolean is passed as a parameter,
the effect type will always be $\boolt \gg \boolt$, similarly for
$\unitt$ and $\funt{T}{U}{V}$. An effect type of $\boolt \gg \unitt$ is not
possible. In general, effect types are kind-preserving.

\begin{figure}
\includegraphics[width=0.49\textwidth]{sm1.pdf}
\caption{\label{fig:sm1} Operational semantics}
\end{figure}

In the case of an object parameter, not all effect types are possible
either. Consider
the object protocol in \figref{fig:sm1}, which we will refer to as $O_1$.
An effect type of $O_1@B \gg O_1@A$ is impossible for this protocol, as there
is no sequence of method calls which would result in a transition from state
$B$ to state $A$. An effect type of $O_1@B \gg O_1@\{B,D\}$ is possible, 
as there is at least one sequence of method calls on an object in
state $B$ to both state $B$ and $D$.
An effect type $O_1@{B,D} \gg O_1@{B,C}$ which has a state set on
the left hand side is possible as there exists a sequence of method calls 
that would take an object in state $B$ to $C$ that would also
take an object in state $D$ to $B$ (specifically, a single method
call to $b$).

In the general case, an effect type concerning an object protocol $O$
of form $O@\overline{S} \gg O@\overline{S_2}$ is
legal if:

\begin{itemize}
\item Given a sequence of method calls that takes an object in state $S \in \overline{S}$
to state $S_2 \in \overline{S_2}$, the same sequence must also take an object
in any other $S' \in \overline{S}$ to a state $S_2' \in \overline{S_2}$.
\item All $S_2 \in \overline{S_2}$ are reachable by at least one sequence
of method calls from an object in a state $S \in \overline{S}$.
\end{itemize}

Checking the legality of an effect type is straightforward: by computing the
intersection of all the states, we can associate with each state in the 
intersection a set of states in the original machine
it was derived from (as shown in \figref{fig:sm1}. 
Should an associated state set contain an end state,
if all other states in the set are also end states then this represents a
``candidate'' end state within the intersection machine. The union of the state 
sets associated with all candidate end states must be equal to the
desired end state set for the effect type to be legal. This would correctly rule 
out the effect type $O@\{B,D\} \gg O@\{B\}$ (there is no sequence of method
calls that would take an instance in either state $B$ or $D$ and result in
the object exclusively being in state $B$) while accepting effect types
$O@\{B,D\} \gg O@\{B,D\}$ (corresponding to no operations occurring) and
$O@\{B,D\} \gg O@\{B,C,D\}$ (corresponding to either no operations occuring
or a single call to $b$).

\subsection{Principal types}

A principal type is one which makes the smallest commitment about 
the structure of a value based on its usage; due to the presence of subtyping 
in TS, a principal type is an upper bound on the type of the value.

\begin{figure}
\includegraphics[width=0.49\textwidth]{principal.pdf}
\caption{\label{fig:principalex}}
\end{figure}

Following from this general definition, a principal object type in TS 
is an object protocol and initial state set which permits all sequences of 
method calls that could be generated by an associated term $t$, and no more. 
In other words,
the language accepted by the state machine is the smallest possible, while
the return types of methods are as general as possible. Consider the example
function $f$ shown in \figref{lst:principalex}. It is clear from the usage
of $y$ that it must be of type $\boolt$ as demanded by the type judgement
of an if-then-else statement, while $x$ is an object of some sort.
We could ascribe any of the object types $O_1@A$, $O_2@A$ and $O_3@A$ to the parameter 
$x$, and after evaluation the object types would be $O_1@A$, $O_2@\{E,H\}$
and $O_2@F$ respectively.
$O_1@A$ is not principal as it will admit many more sequences of method calls 
than strictly required by the body of function $f$, while $O_2@A$ and $O_3@A$ 
are principal.

$O_2$ contains more states than are strictly necessary, i.e. it
is not a {\tt minimal} state machine, while $O_3$ is minimal.
We do not require that a principal type be minimal. If necessary, any
principal state machine we produce can be minimised, so this can be done
on demand where there is a specific need for it. One such need may be when
presenting the type to a user, as a minimal state machine may be easier to 
understand. Regardless, the state machines $O_2$ and $O_3$ can be shown to be 
isomorphic in the language that they accept (starting at state $A$), and that 
language is minimal, so we assert that either are acceptable principal types.

Similarly, it is possible for a principal effect type to have more than
one state in the state set of the input (left) object type. It is
always possible to restructure such an object type to have a single initial
state. It is not, however, possible to have both a single input state and
single output state for a state machine accepting an arbitrary language --- 
consider the language accepted by the regular expression $m^* | n^*$. It is 
necessary for a state machine which accepts this language, and only this 
language, to
have at least three exit states. In general, it is more intuitive to have
a single input state in the effect type but this is not required for a
type to be considered principal.

A principal type we can therefore ascribe to the function $f$
is $\forall T_1 \ldots T_6 . (O_3@A \gg O_3@F, \boolt) \rightarrow T_6$, 
or equivalently 
$\forall T_1 \ldots T_6 . (O_2@A \gg O_2@\{E,H\}, \boolt) \rightarrow T_6$.

There are occasional advantages to not always minimising principal state
machines. The additional structure of $O_2$ retains additional information
in the effect type which can be helpful when interpreting an effect type
when a subtype is passed to the function, as will be described in the next
section.

\subsection{Interpreting effect types}

The typing judgement for a function call permits a passed parameter to
be a subtype of the input portion of the associated effect type.
Determining a precise upper bound for the type of the parameter after the
function call, in particular where the parameter is an object, 
requires some thought.

Consider a function call $f(x)$ where
$f : (O_2@\overline{S_2} \gg O@\overline{S_4}) \rightarrow T$ and
$x : O_1@\overline{S} <: O_2@\overline{S_2}$, where $O_1$ and $O_2$
are the state machines from \figref{fig:principalex}, 
$\overline{S} = \{ A \}$, $S_2 = \{ A \}$ and $S_4 = \{E, H\}$. 
Based on the effect type,
we can assert that the function will have interacted with the $x$ using
one of the possible sequences that would be accepted by $O_2@\overline{S_2}$,
ending with the object in the state set $\overline{S_4}$.

A straightforward but questionable choice would 
be to replace the type of the variable
with that of the output portion of the effect type after the call. However,
this will often result in an unacceptable loss of type information,
reducing the number of valid programs which can be typed within the system.

It is clear that as $\ot{O}{S} \sim \ot{O_2}{S_2}$ there must exist some
$\overline{S_3}$ such that $\ot{O}{S_3} \sim \ot{O_2}{S_4}$.
One possible solution would be to choose
$\overline{S_3} = \{ S | O@S \sim \ot{O_2}{S_4} \}$. This is not specific
enough however --- given the object protocols from \figref{fig:principalex},
$\overline{S_3} = \{ A, B, C, D \}$ as there are no possible method calls
on states $E$ or $H$ of $O_2$.

It is clear that our choice of $S_3$ must relate the set of possible method
call sequences between $\ot{O_2}{S_2}$ and $\ot{O_2}{S_4}$. For this we
define a relation between pairs of object types
$(O@\overline{S}, O_2@\overline{S_2} \lesssim 
(O@\overline{S'}, O_2@\overline{S_2'})$, which we call the
{\it simulation predecessor relation}. The definition of this relation
is given in \figref{fig:simreach}. Essentially, if $O@S \sim O_2@S_2$ and there
exists a sequence of method calls that would transition an object value of type
$O_2@\overline{S_2}$ to $O_2@\overline{S_4}$, then this sequence of method calls
would transition an object in state $O@\overline{S}$ to $O@\overline{S_3}$
such that $O@\overline{S_3} \sim O@\overline{S_4}$. This is abbreviated to
$(O@\overline{S}, O_2@\overline{S_2}) \lesssim (O@\overline{S_3}, O@\overline{S_4})$.

Using this relation, we can choose $\overline{S_3} = 
\bigcup \{ \overline{S'} | 
  (\ot{O}{S} , \ot{O_2}{S_2}) \lesssim (\ot{O}{S'}, \ot{O_2}{S_4})
\}$. The solution using the object protocols from \figref{fig:principalex}
will therefore be $\overline{S_3} = \{ A \}$.

\begin{figure}

\[
\begin{array}{c}
\\
\infer[\lesssim\textsc{-adj}]
  {(O@\overline{S}, O_2@\overline{S_2}) \lesssim (O@\overline{S_3}, O_2@\overline{S_4})}
  {
    \begin{array}{cc}
    O@\overline{S} \sim O_2@\overline{S_2}
    \\
    m : T_i \Rightarrow \overline{S_4} \in O@\overline{S_2}
    \\
    m : T_i' \Rightarrow \overline{S_3} \in O@\overline{S}
    \end{array}
  }
\\\\
\infer[\lesssim\textsc{-refl}]
  {(O@\overline{S}, O_2@\overline{S_2}) \lesssim (O@\overline{S}, O_2@\overline{S_2})}
  {O@\overline{S} \sim O_2@\overline{S_2}}
\\\\
\infer[\lesssim\textsc{-trans}]
  {
  (O@\overline{S}, O_2@\overline{S_2}) \lesssim (O@\overline{S_5}, O_2@\overline{S_6})
  }
  {
  \begin{array}{c}
  (O@\overline{S}, O_2@\overline{S_2}) \lesssim (O@\overline{S_3}, O_2@\overline{S_4})
  \\
  (O@\overline{S_3}, O_2@\overline{S_4}) \lesssim (O@\overline{S_5}, O_2@\overline{S_6})
  \end{array}
  }
\end{array}
\]
\caption{\label{fig:simreach} Definition of the simulation predecessor relation}
\end{figure}

The type system of TS uses a partial function named {\it remap}, defined in
\figref{fig:remaprules} to formalise
this selection of the type of a parameter after a function call. The cases
for Bool, Unit and function types are trivial as these types are immutable,
while for objects the simulation predecessor relation is used.

\begin{figure}
\ottdefnremap

\caption{\label{fig:remaprules} Definition of remap}
\end{figure}

\subsection{Soundness}

The soundness of the type system is stated formally in the following
theorem:

\begin{thm}[Progress and Preservation]
Given a term $t$ such that $\typerule{\Gamma}{t}{T}{\Gamma'}$
with a store $\mu$ such that $\Gamma \vdash \mu$, either $t$ is a value or 
there exist $t'$ and $\mu'$ such that
$\oprule{t}{\mu}{t'}{\mu'}$, with some $\Gamma''$ such that
$\Gamma'' \vdash \mu'$ and $\typerule{\Gamma''}{t'}{T}{\Gamma'}$.
\end{thm}

\begin{proof}
in appendix \appref{app:soundness}.
\end{proof}

\section{Type inference for TS0}

Throughout much of the preceding text we have avoided writing object
types using the full syntax, instead preferring to use diagrams. The full
syntax for object types is verbose and difficult for a user to parse.
Having to write out object types in full would therefore obfuscate the
intent of the code, particularly when there is such intuitively a
strong correlation between the code and the principal object types of the
variables it interacts with. As such, the ability to infer principal object
types is essential for any language which includes typestate, especially
where structural rather than nominal subtyping is employed.

The main source of type verbosity in TS0 is the requirement for
type annotations on function literals. 
Despite there being no need for alias control
annotations, the effect declarations for each parameter are still verbose
and the ability to define type aliases would only help partially.

We therefore desire a general type inference strategy that can infer a
minimal $\Gamma$, $\Gamma'$ and $\alpha$ for a term $t$ such that 
$\typerule{\Gamma}{t}{\alpha}{\Gamma'}$. If $t$ is the body of a function
literal, we can extract the effect types for the named parameters (of which
the free variables of $t$ should be a subset) from $\Gamma$ and $\Gamma'$.

Our approach to type inference is inspired by \cite{Pottier2005} which advocates
separate constraint generation and constraint solving phases for
type inference. Our constraint grammar is shown in \figref{fig:congrammar}.

\begin{figure}
\ottgrammartabular{
\ottC\ottinterrule
\ottctxc\ottinterrule
\otttyc\ottinterrule
\otttinf\ottinterrule
}
\caption{\label{fig:congrammar} Constraint grammar}
\end{figure}

\begin{figure}
\[
\begin{array}{l}

% unit value
\inferlhs{\unitv} \triangleq \alpha = \unitt \cand \gamma = \gamma' 
\\\\

% true value
\inferlhs{\truev} \triangleq \alpha = \boolt \cand \gamma = \gamma' 
\\\\

% false value
\inferlhs{\falsev} \triangleq \alpha = \boolt \cand \gamma = \gamma' 
\\\\

% object value
\inferlhs{[ \overline{S_i \{ \overline{ m_{ij} = (v_{ij}, S_{ij}) } \} }]@S}
\triangleq
\exists \overline{\alpha_{ij}} . \\
\alpha = \{ \overline{ S_i \{ \overline{ m_{ij} : \alpha_{ij} >> S_{ij} } \} }\}@S
\cand
\gamma = \gamma'
\cand \\
\overline{\inferrule{\gamma}{\triangleright v_{ij}}{\alpha_{ij}}{\gamma}}
\\\\

% function value
\inferlhs{\lambda ( \overline{x_i} ) . t} \triangleq
\exists \gamma_r, \gamma_{in}, \gamma_{out}, \overline{\alpha_{ai}, \alpha_{bi}}, \alpha_r . \\
\alpha = (\overline{\alpha_{ai} \gg \alpha_{bi}}) \rightarrow \alpha_r
\cand
\gamma = \gamma'
\cand
\gamma_r = \emptyset
\cand \\
\gamma_{in} = \gamma_r, \overline{x_i : \alpha_{ai}}
\cand
\overline{x_i : \alpha_{bi} \in \gamma_{out}}
\cand
\\
\inferrule{\gamma_{in}}{t}{\alpha_r}{\gamma_{out}}
\\\\

% let binding
\inferlhs{\lett{x}{t'}{t}} \triangleq \exists \gamma_v, \gamma_{b1}, \gamma_{b2}, \alpha' . \\
\gamma_{b1} = \gamma_v, x : \alpha'
\cand
\gamma' = \gamma_{b2} - \{ x \}
\cand \\
\inferrule{\gamma}{t'}{\alpha'}{\gamma_v}
\cand
\inferrule{\gamma_{b1}}{t}{\alpha}{\gamma_{b2}}
\\\\

% sequence
\inferlhs{t ; t'} \triangleq
\exists \gamma'', \alpha' . \\
\inferrule{\gamma}{t}{\alpha'}{\gamma''}
\cand
\inferrule{\gamma''}{t'}{\alpha}{\gamma'}
\\\\

% function call
\inferlhs{x ( \overline{x_i} )} \triangleq
\exists \overline{\alpha_i}, \overline{\alpha_i'}, \overline{\alpha_i''} . \\
\overline{x_i : \alpha_i \in \gamma}
\cand 
x : ( \overline{\alpha_i' \gg \alpha_i''} ) \rightarrow \alpha \in \gamma
\cand
\\
\gamma' = \gamma [ \overline{x_i \mapsto \remapapp{\alpha_i}{\alpha_i'}{\alpha_i''} } ]
\\\\

% method call
\inferlhs{x.m} \triangleq
\exists \alpha_x . \\
x : \alpha_x \in \gamma
\cand 
\cand  \\
\gamma' = \gamma[ x \mapsto \remapapp{\alpha_x}{O@S}{O@S'} ]
\\
\mathrm{where } O = \{ S \{ m : \alpha \Rightarrow S' \} S' \{ \} \}
\\\\

% if-then-else
\inferlhs{\ift{t_c}{t_t}{t_f}} \triangleq
\\
\exists \alpha_c, \alpha_t, \alpha_f, \gamma_c, \gamma_t, \gamma_f . \\
\inferrule{\gamma}{t_c}{\alpha_c}{\gamma_c}
\cand
\inferrule{\gamma_c}{t_t}{\alpha_t}{\gamma_t}
\cand 
\\
\inferrule{\gamma_c}{t_f}{\alpha_f}{\gamma_f}
\cand
\alpha_c = \boolt
\cand
\\
\alpha = \alpha_t \sqcup \alpha_f
\cand
\gamma' = \gamma_t \sqcup \gamma_f
\\\\

\end{array}
\]
\caption{\label{fig:constraintgen} Constraint generation}
\end{figure}

There are three categories of constraints:

\begin{itemize}
\item {\bf Context constraints} --- these define relationships between contexts.
It is important that type changes be correctly observed by the inference
approach and that context constraints expressed in terms of equality, extension,
modification and removal exactly mirror the typing judgements. 
Extension, replacement and removal constraints modify the solution of
another context variable. Following these context variable dependencies,
ultimately every context variable depends upon a {\it root} variable which
is either of form $\gamma = \emptyset$ or $\free{\gamma}$. The root
context is {\it free} when it is the input context to the top-level term being
processed, while
the root is {\it empty} when it is the input context for a function body term.
A free context can be arbitrarily extended with new variables, while an
empty context cannot --- this difference allows the detection of unbound
variable usage within function bodies and the collection of free variables
in the root term.

Additionally, we have {\bf variable usage constraints} of form 
$x : \mathbb{T} \in \gamma$, which are generated for each location in the
program where a variable is used and what type that usage implies. These
are used to generate additional type variable
constraints by matching the asserted type $\mathbb{T}$ with what is known
about the type of $x$ in $\gamma$.

\item {\bf Type expressions constraints} --- these express 
constraints between two or three types: either two types are equal,
or one type is the join of two others. The latter form of type constraint
is generated when dealing with if-then-else terms.
\end{itemize}

Type variables can appear at any position within a type's structure, as shown
in the grammar for $\mathbb{T}$. Objects are treated as a special case, where
we use variables as placeholders for the object definition and state parts.
The states and transitions of the object structure are determined at a later
stage in the type inference process.

A constraint generation expression $\inferlhs{t}$, which relates the term
to an input context variable, a type variable and an output context variable, 
is expanded using the rules shown in \figref{fig:constraintgen}.
The constraint generation process for a root term $t$ begins with
$\exists \gamma, \gamma', \alpha. \free{\gamma} \cand \inferlhs{t}$.

The constraint generation rules correspond to the typing judgements for 
well-typed terms as one would expect:

\begin{itemize}
\item All value terms produce a $\gamma = \gamma'$ constraint as values do
not change the context. Additionally, a type variable constraint of form
$\alpha = \mathbb{T}$ where $\mathbb{T}$ is of the appropriate type to match
the value, with type variables in all positions relating to parameters and return
values.
\item The context variable $\gamma_a$ used as the input context for
function bodies has an empty root $\gamma_r$, to allow for detection of 
used unbound variables within
the body. This matches the use of a context in type inference
composed only of the parameters, and prevents implicit binding.
\item Let-bind terms generates a new context variable $\gamma_v$ for the
output context of the value term. This is extended by $\gamma_{b1}$, the
input context for the body term, with variable $x$ of the type inferred
for the value. The output context of the whole let-bind term removes
$x$ from the output context $\gamma_{b2}$ of the body term.
\item Update terms generate replacement constraints based on the inferred
type $\alpha'$ of the right-hand side of the expression.
\item Sequencing terms relate the output of the left term to the input of the
right term through a new context variable $\gamma''$. The type variable
$\alpha'$ must be generated to capture the type of the left term for
consistency, but it is not needed.
\item Function call terms generate context variable constraints to relate the
usage of the variables to whatever information has been inferred about the
parameter types of the function. All variables $\overline{x_i}$ used as 
parameters in the function call are replaced in the output context with
new type bindings to $\overline{\alpha_i'}$, which represents the effect
of the function.
\item Method call terms generate type variables to capture that $x$ must
be an object, that the method $m$ must exist in the current state $\alpha_s$
and that after the method call $x$ must be in the new state $\alpha_s'$.
\end{itemize}

\subsection{Context constraint solving}

Constraint solving is performed in phases. In the first phase, the context 
constraints
are normalised using the rules in \figref{fig:contextsolve} --- where the set of 
constraints in the premise exists, they are replaced with those in the 
conclusion. Normalisation means that every context variable will have a single 
defining constraint of form $\gamma = \emptyset$, $\free{\gamma}$ or
$\gamma = \gamma', \overline{x_i : \mathbb{T}_i}$, where $\gamma'$ is
either empty or free. A context $\Gamma$ for type checking can be easily
constructed from constraints in this form, as once the type variable
constraints are solved we can say $\Gamma = \overline{x_i : T_i}$ for
every $\gamma = \gamma', \overline{x_i : \mathbb{T}_i}$ where $T_i$ is the
solved form of $\mathbb{T}_i$.

The context normalisation rules do the following:

\begin{itemize}
\item \textsc{ceq} --- Where $\gamma$ is directly equivalent
to $\gamma'$, which is defined in terms of extension of
$\gamma''$, rewrite the constraints such that
$\gamma$ directly copies the extension of $\gamma''$.
\item \textsc{crem} --- Where $\gamma$ removes a variable $x$ from $\gamma'$,
which is defined in terms of extension of $\gamma''$, rewrite the
constraints such that $\gamma$ copies the extension of $\gamma''$ with
the variable $x$ removed from the extension. There is no need for an
error-producing variant of \textsc{crem}, where the variable $x$ is missing, 
as context removal constraints are always generated in a pair with a context 
extension constraint for let binding.
\item \textsc{cext} --- Where $\gamma$ extends $\gamma'$ with
a variable set $\overline{y_j}$, and $\gamma'$ is defined in terms of extension of
$\gamma''$ with a disjoint set of variables in its extension set, rewrite the
constraints such that $\gamma$'s extension set includes both the $y_j$
definitions and $x_j$ definitions.
\item $\textsc{cext}_2$ --- Where $\gamma$ extends $\gamma'$ with
a new variable $x$, and $\gamma'$ is defined in terms of extension of
some $\gamma''$ and the extension set already contains $x$, this
represents an attempt to redeclare $x$. As such, generate an error
$\mathbf{bound}\:x$, terminating the constraint solving process.
\item \textsc{crep} --- where $\gamma$ replaces the type of some
variable $x$ in $\gamma'$, where $\gamma$ is defined in terms of extension
of $\gamma''$ and $x$ is in the extension set, rewrite the constraints
such that $\gamma$ copies the extension set of $\gamma'$ with $x$
replaced as specified.
\item $\textsc{crep}_2$ --- where $\gamma$ replaces variable $x$ in
$\gamma'$ which is defined in terms of extension of $\gamma''$
which is free and $x$ is not in the extension set, $x$ is a
free variable in the term being processed. To introduce $x$ in such a way that
it will be included in all contexts, $\gamma''$ is redefined in
terms of extension of a new context $\gamma_r$ with variable $x$
in the extension set mapped to a new type variable $\alpha$. Following
this, \textsc{cext} can be applied to $\gamma'$, then \textsc{crep} can
be applied to $\gamma$.
\item $\textsc{crep}_3$ --- where $\gamma$ replaces variable $x$ in
$\gamma'$ which is defined in terms of an empty root $\gamma''$,
and $x$ is not in $\gamma$'s extension set, this indicates the usage
of an unbound variable within a function body. As such, generate an
error $\mathbf{unbound}\:x$ indicating that x is unbound within this
context, terminating the contraint solving process.
\item \textsc{cvar} --- where a variable constraint for $x$ exists
for context variable $\gamma$, which is defined in terms of
extension of $\gamma_r$ which is free, and $x$ is not in the extension
set of $\gamma$, $x$ is a free variable in the term being processed. 
Similar to $\textsc{crep}_2$, redefine $\gamma_r$
in terms of extension of a new root $\gamma_r'$, with $x$ in
the extension set with the observed type $\mathbb{T}$.
\item $\textsc{cvar}_2$ --- where a variable constraint for $x$ exists
for context variable $\gamma$, which is defined in terms of $\gamma'$
which is empty, and $x$ is not in the extension set of $\gamma$, 
$x$ is an unbound variable within a function body.
\end{itemize}

The rules are repeatedly applied to any matching sets of constraints
until no further rewriting is possible, producing either an error
or a set of context constraints which are fully normalised, and
can be converted into contexts for type checking easily.

The way in which the context constraints are generated guarantees
that the dependencies between contexts form a directed acyclic graph. For
this reason, we believe that that the first phase will always terminate,
but this has not yet been proven.

\subsection{Variable usage \& type variable constraint solving}

In the second phase, additional type variable 
constraints are generated by following the rule in \figref{fig:tvargen}, which
compares the observed type from the usage of a variable to the type
known in the context. Repeated application of this rule until no more matches
are found produces a constraint set composed exclusively of normalised
context variable constraints, type variable constraints and protocol
constraints. A first order unification algorithm can then be used to solve
the system of type variable constraints, producing a most general unifier
for the type variables if unification is possible.

First order unification will identify a set of type variables (and their synonyms)
which represent distinct object types. For each distinct object type, there
will also be a set of type variables which represent distinct states for
the object type. By matching these variables to the protocol constraints
within the constraint set, we can generate the state machine for each
object type in a straightforward manner. To produce a minimal description of 
the object protocol, this state machine can be minimised by constructing a 
bisimulation relation between the states within each object and collapsing 
equivalent states.

The typing rules of TS0 do not permit polymorphism, so
if the most general unifier does not map all type variables for
variables that occur in the input and output contexts of the term being
processed, then the term has no valid type in the language. Otherwise,
substituting the fully explicit type inferred for each type variable that
occurs in the inferred input and output context variables, as well as the
type variable for the input term, should produce a valid typing judgement
where the domain of the input context is the set of free variables in
the term. This has not yet been proven.

\begin{figure}

\begin{center}
\begin{tabular}{c}
	\infer[\textsc{ceq}]
	{\gamma = \gamma'', \overline{x_i : \mathbb{T}_i} \cand \gamma' = \gamma'', \overline{x_i : \mathbb{T}_i}}
	{\gamma = \gamma' \cand \gamma' = \gamma'', \overline{x_i : \mathbb{T}_i}}
\\\\
	\infer[\textsc{crem}]
	{
	\gamma = \gamma'', \overline{x_i : \mathbb{T}_i} \cand \gamma' = \gamma'', \overline{x_i : \mathbb{T}_i}, x : \mathbb{T}
	}
	{
	\gamma = \gamma' - \{x\} \cand \gamma' = \gamma'', \overline{x_i : \mathbb{T}_i}, x : \mathbb{T}
	}
\\\\
	\infer[\textsc{cext}]
	{\gamma = \gamma'', \overline{x_i : \mathbb{T}_i}, \overline{y_j : \mathbb{T}_j} \cand \gamma' = \gamma'', \overline{x_i : \mathbb{T}_i}}
	{
		\gamma = \gamma', \overline{y_j : \mathbb{T}_j} \cand \gamma' = \gamma'', \overline{x_i : \mathbb{T}_i}
		& \overline{y_j} \cap \overline{x_i} = \emptyset
	}
\\\\
	\infer[\textsc{cext\_2}]
	{\mathbf{bound}\:x}
	{
		\gamma = \gamma', \overline{y_j : \mathbb{T}_j}, x : \mathbb{T} \cand \gamma' = \gamma'', \overline{x_i : \mathbb{T}_i}
		& x \in \overline{x_i}
	}
\\\\
	\infer[\textsc{crep}]
	{\gamma = \gamma'', \overline{x_i : \mathbb{T}_i}, x : \mathbb{T} \cand
	\gamma' = \gamma'', \overline{x_i : \mathbb{T}_i}, x : \mathbb{T}'}
	{\gamma = \gamma'[x \mapsto \mathbb{T}] \cand
	\gamma' = \gamma'', \overline{x_i : \mathbb{T}_i}, x : \mathbb{T}'}
\\\\
	\infer[\textsc{crep\_2}]
	{
	\begin{array}{c}
	\exists \alpha, \gamma_r . 
	\gamma = \gamma'[x \mapsto \mathbb{T}] \cand
	\gamma' = \gamma'', \overline{x_i : \mathbb{T}_i} \cand
	\\
	\gamma'' = \gamma_r, x : \alpha \cand
	\free{\gamma_r}
	\end{array}
	}
	{
	\begin{array}{c}
	\begin{array}{ccc}
	x \notin \overline{x_i}
	&
	\fresh{\gamma_r}
	&
	\fresh{\alpha}
	\end{array}
	\\
	\gamma = \gamma'[x \mapsto \mathbb{T}] \cand
	\gamma' = \gamma'', \overline{x_i : \mathbb{T}_i} \cand
	\free{\gamma''}
	\end{array}
	}
\\\\
	\infer[\textsc{crep\_3}]
	{\mathbf{unbound}\:x}
	{
	\begin{array}{c}
	x \notin \overline{x_i}
	\\
	\gamma = \gamma'[x \mapsto \mathbb{T}] \cand
	\gamma' = \gamma'', \overline{x_i : \mathbb{T}_i} \cand
	\gamma'' = \emptyset
	\end{array}
	}
\\\\
	\infer[\textsc{cvar}]
	{\exists \gamma_r' . 
	\gamma = \gamma_r, \overline{x_i : \mathbb{T}_i} \cand
	\gamma_r = \gamma_r', x : \mathbb{T} \cand
	\free{\gamma_r'}
	}
	{
	\begin{array}{c}
	\begin{array}{cc}
	x \notin \overline{x_i}
	&
	\fresh{\gamma_r'}
	\end{array}
	\\
	\gamma = \gamma_r, \overline{x_i : \mathbb{T}_i} \cand 
	x : \mathbb{T} \in \gamma \cand
	\free{\gamma_r}
	\end{array}
	}
\\\\
	\infer[\textsc{cvar\_2}]
	{\mathbf{unbound}\:x}
	{x : \mathbb{T} \in \gamma \cand \gamma = \gamma', \overline{x_i : \mathbb{T}_i} \cand \gamma' = \emptyset
	&
	x \notin \overline{x_i}}
\end{tabular}
\end{center}
\caption{\label{fig:contextsolve} Context normalisation}
\end{figure}

\begin{figure}
\[
	\infer[\textsc{tvar}]
	{\mathbb{T} = \mathbb{T}' \cand \gamma = \gamma', \overline{x_i : \mathbb{T}_i}, x : \mathbb{T}'}
	{x : \mathbb{T} \in \gamma \cand \gamma = \gamma', \overline{x_i : \mathbb{T}_i}, x : \mathbb{T}'}
\]
\caption{\label{fig:tvargen} Refining type variable constraints}
\end{figure}

\subsection{Implementation}

A Scala implementation of a type checker for TS0 which uses the type inference
technique described can be found at \url{http://github.com/iainmcgin/ts0-lang}.
A topological order is built from the dependency graph between context variables
to efficiently normalise the context constraints. The first order unification 
algorithm presented in \cite{Martelli1982} is used to solve the type variable 
constraints, allowing the effects for the parameters to be extracted from
the context variables.

\section{Conclusion}

TS0 is a small, simple language that has proven useful
for the study of the inference of object protocols, and the theoretical work
done to date holds promise for future expansion and interesting results. 
To our knowledge, this is also the first attempt to study the inference
of generalised typestate in an object oriented context, rather than
more specific inference problems such as session type inference
\cite{Collingbourne2010}. The
constraint generation approach closely aligns with the typing judgements,
which we believe makes the approach intuitive.

\appendix

\section{Soundness}

\begin{thm}[Progress and Preservation]
Given a term $t$ such that $\typerule{\Gamma}{t}{T}{\Gamma'}$
with a store $\mu$ such that $\Gamma \vdash \mu$, either $t$ is a value or 
there exist $t'$ and $\mu'$ such that
$\oprule{t}{\mu}{t'}{\mu'}$, with some $\Gamma''$ such that
$\Gamma'' \vdash \mu'$ and $\typerule{\Gamma''}{t'}{T}{\Gamma'}$.
\end{thm}

\begin{proof}
by induction on the structure of $t$. Assume $t$ is not a value. It is
therefore of one of the following forms:

\begin{itemize}
\item $t = \lett{x}{t''}{t'}$. 
It follows by rule \tlet that there exist $\Gamma_1, T', T''$
such that
$\typerule{\Gamma}{t''}{T'}{\Gamma_1}$ and that
$\typerule{\Gamma_1, x : T'}{t'}{T}{\Gamma', x : T''}$. There are two
possibilities:

  \begin{itemize}
  \item If $t''$ is some value $v$, then by \lemref{lem:valuectx} 
  we have that $\Gamma = \Gamma_1$ and reduction can occur
  by \rletv, such that 
  $\oprule{\lett{x}{v}{t'}}{\mu}{t'}{\mu [ x \mapsto v ]}$.
  Let $\mu' = \mu [ x \mapsto v ]$. Let $\Gamma'' = \Gamma, x : T'$.
  It follows that that $\Gamma'' \vdash \mu'$, with
  $\typerule{\Gamma''}{t'}{T}{\Gamma', x : T'}$.
  The requirements of rule \tdrop are satisfied by this, therefore we can 
  type $t'$ such that $\typerule{\Gamma''}{t'}{T}{\Gamma'}$.

  \item If $t''$ is a term, 
  by induction we have that $\oprule{t''}{\mu}{t'''}{\mu'}$, with
  some $\Gamma'' \vdash \mu'$ and $\typerule{\Gamma''}{t'''}{T'}{\Gamma_1}$.
  This satisfies the requirements of \rlett, meaning $t$ itself can reduce
  such that $\oprule{\lett{x}{t''}{t'}}{\mu}{\lett{x}{t'''}{t'}}{\mu'}$.
  Additionally, the conditions of \tlet are satisfied such that
  $\typerule{\Gamma''}{\lett{x}{t'''}{t'}}{T}{\Gamma'}$.

  \end{itemize}

\item $t = t_l ; t_r$. It follows by rule
\tseq that $\typerule{\Gamma}{t_l}{T_l}{\Gamma_{mid}}$ and
$\typerule{\Gamma_{mid}}{t_r}{T}{\Gamma'}$.

  \begin{itemize}
  \item If $t_l$ is some value $v$, then
  reduction can occur by \\ 
  \rseqlv such that
  $t_l; t_r \mid \mu \longrightarrow t_r \mu$. By \lemref{lem:valuectx}, 
  $\Gamma'' = \Gamma = \Gamma_{mid}$
  and $\mu' = \mu$, therefore $\Gamma'' \vdash \mu'$ and 
  $\typerule{\Gamma''}{t_r}{T}{\Gamma'}$.

  \item If $t_l$ is a term, by induction $t_l \mid \mu \longrightarrow t_l' \mid \mu'$
  with some $\Gamma'' \vdash \mu'$ and
  $\typerule{\Gamma''}{t_l'}{T_l}{\Gamma_{mid}}$.
  This satisfies the requirements of \rseqlt
  such that $t_l; t_r \mid \mu \longrightarrow t_l'; t_r \mid \mu'$.
  Additionally, the requirements of \tseq are satisfied such that
  $\typerule{\Gamma''}{t_l'; t_r}{T}{\Gamma'}$.

  \end{itemize}

\item $t = x ( \overline{x_i} )$. It follows by rule \tfunc that
$\Gamma = \Gamma_1, \overline{x_i : T_i}$ 
with $\Gamma_1(x) = \funt{U_i}{V_i}{T}$
such that $\overline{T_i <: U_i}$ 
and that $\Gamma' = \Gamma_1, \overline{x_i : T_i'}$
where $T_i' = \mathbf{remap} ( T_i, U_i >> V_i )$.

As $\Gamma \vdash \mu$, 
it follows that $\mu(x) = \funv{y_i}{U_i'}{V_i'}{t'}$
where $\typerule{\overline{y_i : U_i'}}{t'}{T'}{\overline{y_i : V_i'}}$,
with $\funt{U_i'}{V_i'}{T'} <: \funt{U_i}{V_i}{T}$. It
follows by \subfn that for each $i$, $U_i <: U_i'$, $V_i' <: V_i$
and $T' <: T$.
Additionally for each $x_i$, $\mu(x_i) = v_i$ such that 
$\typerule{\emptyset}{v_i}{T_i''}{\emptyset}$ 
where $T_i'' <: T_i$.

Let $t_{sub} = \subst{x_i}{y_i}{t'}$.
The term $t$ can be reduced by \rfunc such that
$x ( \overline{x_i} ) \mid \mu \longrightarrow t_{sub} \mid \mu$.
By the substitution lemma, 
$\typerule{\overline{x_i : U_i'}}{t_{sub}}{T'}{\overline{x_i : V_i'}}$.
By lemma \lemref{lem:upgrading}, it follows that
for each $i$, there exists a $V_i'' = \remapapp{T_i''}{U_i'}{V_i'}$
such that 
$\typerule{\overline{x_i : U_i}}{T_{sub}}{T'}{\overline{x_i : V_i''}}$.
By \lemref{lem:remap} for each $i$ that $V_i'' <: V_i'$.

Therefore, by given that $T_i <: U_i$ and that $V_i'' <: V_i' <: V_i$
and that $T' <: T$ we can apply \tnarrow, \twiden and \tsub
such that $\typerule{\overline{x_i : T_i}}{t'}{T}{\overline{x_i : V_i}}$.

Let $\mu' = \mu$. Let $\Gamma'' = \Gamma$. It follows that 
$\Gamma'' \vdash \mu'$. By applying the weakening lemma, we can type 
$t_{sub}$ such that $\typerule{\Gamma''}{t''}{T}{\Gamma'}$.

\item $t = x.m$. It follows by rule \tmethc that
$\Gamma = \Gamma_1, x : O@\overline{S}$ and that 
$\Gamma' = \Gamma_1, x : O@\overline{S'}$, with
$\overline{S'} = \{ S_n | m : T_r \Rightarrow S_n \in O@S, S \in \overline{S} \}$.
As
$\Gamma \vdash \mu$, we must have $\mu = \mu_1, x \mapsto o'@S_c$ with
$\Gamma_1 \vdash \mu_1$ and where
$\typerule{\emptyset}{o'@S_c}{O'@S_c}{\emptyset}$ such that
$O'@S_c <: O@\overline{S}$.

By the definition of subtyping for object types,
we must have that
$o' = [ S' \{ m = (v,S_n) ; ... \} ... ]$, where $O'@S_n <: O@\overline{S'}$. 
Where the value $v$ is concerned, $\typerule{\emptyset}{v}{T'}{\emptyset}$
for some $T'$ such that $T' <: T = \bigsqcup{ \{ T_r | m : T_r \Rightarrow S_n \in O@S, S \in \overline{S} \} }$.

Therefore reduction can occur by \rmethc such that
$x.m \mid \mu_1, x \mapsto o@S \longrightarrow v \mid \mu_1, x \mapsto o@S_n$.
Let $\mu' = \mu_1, x \mapsto o@S'$. Let $\Gamma'' = \Gamma'$. It follows that
$\Gamma'' \vdash \mu'$, as $O'@S_n <: O@\overline{S'}$. 
By application of the weakening lemma and rule \tsub, it follows that
$\typerule{\Gamma''}{v}{T}{\Gamma'}$.

\item $t = \ift{t_c}{t_t}{t_f}$. It follows by rule \tif that there exists
$\Gamma_1$, $\Gamma_2$, $\Gamma_3$, $T_t$ and $T_f$ such that
$\typerule{\Gamma}{t_c}{\boolt}{\Gamma_1}$ and that
$\typerule{\Gamma_1}{t_t}{T_t}{\Gamma_2}$ and that
$\typerule{\Gamma_2}{t_f}{T_f}{\Gamma_3}$ with
$\Gamma' = \Gamma_2 \sqcup \Gamma_3$ and $T = T_t \sqcup T_f$. There are
three possibilities for reduction:

  \begin{itemize}
  \item $t_c = \truev$. By \lemref{lem:valuectx} we have that
  $\Gamma = \Gamma_1$. Reduction can occur by \riftrue such that
  $\oprule{t}{\mu}{t_t}{\mu}$. By \tsub we have that
  $\typerule{\Gamma_1}{t_t}{T}{\Gamma_2}$ as $T_t <: T$. Additionally, for
  all $x \in \Gamma_2$, $\Gamma_2(x) <: \Gamma'(x)$. Therefore by
  repeatedly applying \twiden for all $x$, we have that
  $\typerule{\Gamma_1}{t_t}{T}{\Gamma'}$.
  Let $\Gamma'' = \Gamma$, $\mu' = \mu$ and 
  $t' = t_t$. Therefore, $\typerule{\Gamma''}{t'}{T}{\Gamma'}$ with
  $\Gamma'' \vdash \mu'$.

  \item $t_c = \falsev$. By \lemref{lem:valuectx} we have that
  $\Gamma = \Gamma_1$. Reduction can occur by \riffalse such that
  $\oprule{t}{\mu}{t_f}{\mu}$. By \tsub we hhave that
  $\typerule{\Gamma_1}{t_f}{T}{\Gamma_3}$ as $T_f <: T$. Additionally, for
  all $x \in \Gamma_3$, $\Gamma_3(x) <: \Gamma'(x)$. Therefore by
  repeatedly applying \twiden for all $x$, we have that
  $\typerule{\Gamma_1}{t_f}{T}{\Gamma'}$. Let $\Gamma'' = \Gamma$,
  $\mu' = \mu$ and $t' = t_f$. Therefore, 
  $\typerule{\Gamma''}{t'}{T}{\Gamma'}$ with $\Gamma'' \vdash \mu'$.

  \item $t_c$ is not a value. By induction, there exists $t_c'$,
  $\mu'$, $\Gamma''$ such that 
  $\oprule{t_c}{\mu}{t_c'}{\mu'}$ and 
  $\typerule{\Gamma''}{t_c'}{\boolt}{\Gamma_1}$ with $\Gamma'' \vdash \mu'$. 
  Reduction of $t$ can therefore occur by \rifexpr such that
  $\oprule{t}{\mu}{\ift{t_c'}{t_t}{t_f}}{\mu'}$. It follows
  by \tif that $\typerule{\Gamma''}{\ift{t_c'}{t_t}{t_f}}{T}{\Gamma'}$.
  \end{itemize}

\end{itemize}
\end{proof}

\section{Supporting lemmas}
\label{app:proofs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% lem:valuectx %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{lem}
\label{lem:valuectx}
If $v$ is a value and $\typerule{\Gamma}{v}{T}{\Gamma'}$, then $\Gamma = \Gamma'$.
\end{lem}
\begin{proof}
Directly from the typing judgements. $v$ is either {\tt unit}, {\tt true},
{\tt false}, a function literal or an object literal.
Correspondingly, by rules \tunit,
\ttrue, \tfalse, 
\tfundef and \tobj, $\Gamma = \Gamma'$.
\end{proof}

\begin{lem}[Object state sets]
Given an object state set $\overline{S}$ and a state $S'$, if
$S' \in \overline{S}$ then $O@S' <: O@\overline{S}$.
\end{lem}
\begin{proof}
$O@\overline{S}$ is defined as equivalent to
$\bigsqcup_{S \in \overline{S}}{O@S}$. Given that $S \in \overline{S}$,
we must have that $\overline{S} = \overline{S'} \cup \{ S \}$ for
some $\overline{S'}$.
it follows that $O@\overline{S} = O@S' \sqcup \left( 
\bigsqcup_{S \in \overline{S'}}{O@S} \right)$.
By the definition of join, 
$O@S' <: O@S' \sqcup \bigsqcup_{S \in \overline{S}}{O@S}$.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% lem:ssetinc  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{lem}
\label{lem:ssetinc}
Given an object protocol $O = \left\{ \overline{ S_i \{ ... \} } \right\}$ and two state sets 
$\overline{S} \subseteq \overline{S'} \subseteq \overline{S_i}$, it follows
that $O@\overline{S} <: O@\overline{S'}$.
\end{lem}
\begin{proof}
From the definition of state sets, we have that 
$O@\overline{S} = \bigsqcup_{S \in \overline{S}}{O@S}$ and that
$O@\overline{S'} = \bigsqcup_{S \in \overline{S'}}{O@S}$.
As $\overline{S} \subseteq \overline{S'}$ there exists an
$\overline{S''}$ such that $\overline{S'} = \overline{S} \cup \overline{S''}$
where $\overline{S} \cap \overline{S''} = \emptyset$. Therefore
$O@\overline{S'} = \left(\bigsqcup_{S \in \overline{S}}{O@S}\right) \sqcup \left(\bigsqcup_{S' \in \overline{S'} - \overline{S}}{O@S'}\right)$.

As $A <: A \sqcup B$ for all $A$ and $B$ where $A \sqcup B$ is defined, 
it follows that $O@\overline{S} <: O@\overline{S'}$.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% lem:remap %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{lem}
\label{lem:remap}
Given types $T$, $U$, and $V$, such that
$\remapapp{T}{U}{V}$ is defined, it follows that $\remapapp{T}{U}{V} <: V$.
\end{lem}
\begin{proof}
By case analysis of V.

\begin{itemize}
\item $V = \unitt$ or $V = \boolt$. By either \remapunit or \remapbool
it must be the case that $U = V$. It must also be the case that
$T = U$, as $\unitt$ and $\boolt$ do not have any subtypes other than
themselves. Therefore $V = \remapapp{T}{U}{V} <: V$ by \subrefl.

\item $V = \funv{V_i}{V_i'}{V'}$ for some $\overline{V_i}$, $\overline{V_i'}$
and $V'$. By \remapfunc, $T <: U$ and $T = \remapapp{T}{U}{V}$. Therefore,
$T <: V$.

\item $V = O@\overline{S'}$ for some $O$ and $\overline{S'}$. By
\remapobj, $O'@\overline{S'''} = \remapapp{T}{U}{V}$ where 
$T = O'@\overline{S''}$ and $U = O@\overline{S}$ and
$\overline{S'''} = \{ S | O'@S \sim O@\overline{S'}, S \in O' \}$ 
for some $O'$, $\overline{S}$ and $\overline{S''}$. By \subobj,
$T = O'@\overline{S'''} <: O@\overline{S'} = V$ as for each $S \in \overline{S'''}$,
$O'@S <: O@\overline{S'}$.
\end{itemize}
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% lem:remapcv %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{lem}[Remap is covariant]
\label{lem:remapcv}
Given types $T, T', U$ and $V$ such that $T <: T'$, 
$\remapapp{T}{U}{V} <: \remapapp{T'}{U}{V}$.
\end{lem}

\begin{lem}
\label{lem:remapeff}
Given types $T$, $U$ and $V$ such that $U <: V$ and $T = \remapapp{U}{V}{V}$, 
it follows that $U <: T$. \\
(TODO: object portion needs reworked due to new definition of remap)
\end{lem}
\begin{proof}
By case analysis of the type $V$.

\begin{itemize}
\item $V = \unitt$ or $V = \boolt$. 
It must be the case that $V = U$ also, as $\unitt$ and $\boolt$
have no subtypes other than themselves by rule \subrefl.
Therefore, by rule \remapunit or \remapbool, $T = V$,
and $U <: T$ by rule \subrefl.

\item $V = \funv{T_i}{T_i'}{T_r}$ for some $\overline{T_i}$, $\overline{T_i'}$
and $T_r$. As $U <: V$, it follows by rule \remapfunc that
$T = U$, and $U <: T$ by \subrefl.

\item $V = O@\overline{S}$ for some $O$ and $\overline{S}$. 
It follows that $U = O'@\overline{S'}$ 
for some $O'$ and $\overline{S'}$ 
such that $O'@\overline{S'} \sim O@\overline{S}$ 
by rule \subobj. 
By rule \remapobj, it follows that $T = O'@\overline{S''}$ 
such that $\overline{S''} = \{ S | O'@S \sim O@\overline{S}, S \in O' \}$.
By \lemref{lem:ssetinc} it follows that 
$O'@\overline{S'} <: O'@\overline{S''}$.

\end{itemize}
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% lem:remapsame %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{cor}
\label{lem:remapsame}
Given a type $T$, $T <: \remapapp{T}{T}{T}$. \\
(TODO: in fact, with the new definition of remap it should be possible to
state that $T = \remapapp{T}{T}{T}$).
\end{cor}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% lem:upgrade %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{lem}[remap]
\label{lem:upgrade}
If a term $t$ can be typed such that 
$\typerule{\Gamma, x : T}{t}{U}{\Gamma', x : V}$, and if
$T' <: T$, then there exists a $V' <: \remapapp{T'}{T}{V}$ 
such that $\typerule{\Gamma, x : T'}{t}{U}{\Gamma', x : V'}$.
\end{lem}
\begin{proof}
by induction on the structure of $t$.
The term $t$ is of one of the following forms:

\begin{itemize}
\item $t$ is a value, so by \lemref{lem:valuectx} it follows that
$\Gamma = \Gamma'$ and $T = V$. 
The environment can be arbitrary in the typing rules for values,
therefore we can type $t$ such that 
$\typerule{\Gamma, x : T'}{t}{U}{\Gamma', x : T'}$.
Let $V' = T'$.
By \lemref{lem:remapeff}, we have that $V' <: \remapapp{T'}{T}{V}$.
Therefore, $\typerule{\Gamma, x : T'}{t}{U}{\Gamma', x : V'}$.

\item $t = \lett{y}{t'}{t''}$. It follows from the rule \tlet that
$x \neq y$ (as variables cannot be rebound) and that there exists 
$\Gamma_1, \Gamma_2, T_2, T_y, T_y'$ such that
$\typerule{\Gamma, x : T}{t'}{T_y}{\Gamma_1, x : T_2}$ and that
$\typerule{\Gamma_1, x : T_2, y : T_y}{t''}{U}{\Gamma', x : V, y : T_y'}$.
By induction, it follows that there exists a type $V_2 <: \remapapp{T'}{T}{T_2}$ such that
$\typerule{\Gamma, x : T'}{t'}{T_y}{\Gamma_1, x : V_2}$.
By \lemref{lem:remap} we have that $V_2 <: T_2$.
Therefore it also follows by induction that
there exists a type $V_3 <: \remapapp{V_2}{T_2}{V}$
such that
$\typerule{\Gamma_1, x : V_2, y : T_y}{t''}{U}{\Gamma', x : T_3, y : V_3}$.
Let $V' = V_3$. The rules of \tlet are satisfied such that
$\typerule{\Gamma, x : T'}{t}{U}{\Gamma', x : V'}$.

\item $t = y ( \overline{z_i} )$. 
It follows from the rule \tfunc that there exists
$\Gamma_1, \overline{T_i}, \overline{U_i}, \overline{V_i}, \overline{T_i'}$ 
such that 
$\Gamma, x : T = \Gamma_1, y : \funt{U_i}{V_i}{U}, \overline{z_i : T_i}$
and that
$\Gamma', x : V = \Gamma_1, y : \funt{U_i}{V_i}{U}, \overline{z_i : T_i'}$
such that for each $i$, $T_i <: U_i$, $T_i' <: V_i$ 
and $U <: U'$.

There are three cases to consider:
\begin{itemize}
\item $x \neq y$ and $x \notin \overline{x_i}$, meaning that
$\Gamma_1 = \Gamma_2, x : T$ for some $\Gamma_2$, that
$\Gamma = \Gamma_2, y : \funt{U_i}{V_i}{U}, \overline{z_i : T_i}$
and that $\Gamma' = \Gamma_2, y : \funt{U_i}{V_i}{U}, \overline{z_i : T_i'}$.
Additionally, $T = V$. 
It follows that $t$ can be typed without $x$
such that $\typerule{\Gamma}{t}{U}{\Gamma'}$. Let $V' = T'$.
By the weakening lemma, $\typerule{\Gamma, x : T'}{t}{U}{\Gamma', x : V'}$.
By \lemref{lem:remapeff}, $V' <: \remapapp{T'}{T}{V}$.

\item $x = y$, meaning $\Gamma = \Gamma_1, \overline{z_i : T_i}$ and
$\Gamma' = \Gamma_1, \overline{z_i : T_i'}$. 
Additionally, $T = V = \funt{U_i}{V_i}{U}$.
$T'$ must therefore be of form $\funt{U_i'}{V_i'}{U'}$ 
such that for each $i$, $U_i <: U_i'$, $V_i' <: V_i$, $U' <: U$ by the
definition of subtyping. Let $V' = T'$.
By \lemref{lem:remapeff}, $V' <: \remapapp{T'}{T}{V}$.
It follows by rule \tfunc, and \tsub to replace $U'$ with $U$, that
$\typerule{\Gamma, x : T'}{t}{U}{\Gamma', x : V'}$.

\item $x \in \overline{z_i}$, meaning that $T = T_i$ and 
$V = T_i' = \remapapp{T}{U_i}{V_i}$ for some $i$.
Let $V' = \remapapp{T'}{U_i}{V_i}$. 
By \lemref{lem:remapcv}, $V' <: V$.
As $T' <: T$, we can type $t$ by rule \tfunc such that
$\typerule{\Gamma, x : T'}{t}{U}{\Gamma, x : V'}$.

\end{itemize}

\item $t = y.m$. By the rule \tmethc it follows that
$\Gamma, x : T = \Gamma_1, y : O@\overline{S}$
and that
$\Gamma', x : V = \Gamma_1, y : O@\overline{S'}$
for some $\Gamma_1, O, \overline{S}$ and $\overline{S'}$
where
$\overline{S'} = \{ S_n | m : T_r \Rightarrow S_n \in O@S, S \in \overline{S} \}$.
Additionally, 
$U = \bigsqcup{ \{ T_r | m : T_r \Rightarrow S_n \in O@S, S \in \overline{S} \} }$.
There are two cases to consider:

\begin{itemize}
\item $x = y$. It follows that $\Gamma = \Gamma_1 = \Gamma'$ and
that $T = O@\overline{S}, V = O@\overline{S'}$.
$T'$ must be of form $O'@\overline{S}''$ such that
$O'@\overline{S}'' \sim O@\overline{S}$ by rule \subobj.

Let $V' = O'@\overline{S}''' = \remapapp{T'}{T}{V}$
where
$S''' = \{ S_n' | m : T_r' \Rightarrow S_n' \in O'@S'', S'' \in \overline{S}''\}$.
Let $U' = \bigsqcup{ \{ T_r' | m : T_r' \Rightarrow S_n' \in O'@S'', S'' \in \overline{S}'' \} }$.
As $O'@\overline{S}'' \sim O@\overline{S}$, it follows that
$O'@\overline{S}''' \sim O@\overline{S}'$.
Therefore, $U' <: U$ and $V' <: V$. By the rule \tmethc, and \tsub
to replace $U'$ with $U$, it follows that
$\typerule{\Gamma, x : T'}{t}{U}{\Gamma', x : V'}$.

\item $x \neq y$. It follows that $\Gamma = \Gamma'', y : O@\overline{S}$
and that $\Gamma' = \Gamma'', y : O@\overline{S'}$
for some $\Gamma''$, $O$ and $\overline{S}$, and that
$T = V$.
By rule \tmethc we have that $\typerule{\Gamma}{t}{U}{\Gamma'}$.
Let $V' = T'$. By \lemref{lem:remapeff}, $V' <: \remapapp{T'}{T}{V}$.
By the weakening lemma, 
$\typerule{\Gamma, x : T'}{t}{U}{\Gamma', x : V'}$.

\end{itemize}

\item $t = t' ; t''$. It follows by rule \tlet that there exists
$\Gamma'', V_{mid}, U'$ 
such that $\typerule{\Gamma, x : T}{t'}{U'}{\Gamma'', x : V_{mid}}$
and that $\typerule{\Gamma'', x : V_{mid}}{t''}{U}{\Gamma', x : V}$.
By induction, it follows that there exists $V_{mid}' <: \remapapp{T'}{T}{V_{mid}}$ 
such that $\typerule{\Gamma, x : T'}{t'}{U'}{\Gamma'', x : V_{mid}'}$.
By \lemref{lem:remap} we have that $V_{mid}' <: V_{mid}$.
There it also follows by induction that 
there exists $V_{end} <: \remapapp{V_{mid}'}{V_{mid}}{V}$ such that
$\typerule{\Gamma'', x : V_{mid}'}{t''}{U}{\Gamma'', x : V_{end}}$.
Let $V' = V_{end}$. By rule \tlet it follows that
$\typerule{\Gamma, x : T'}{t' ; t''}{U}{\Gamma'', x : V'}$.

\item $t = \ift{t_c}{t_t}{t_f}$. It follows from rule \tif that there
exists $\Gamma_1$, $\Gamma_2$, $\Gamma_3$, $U_t$, $U_f$, $V_c$, $V_t$ and
$V_f$
such that
$\typerule{\Gamma, x : T}{t_c}{\boolt}{\Gamma_1, x : V_c}$ and that
$\typerule{\Gamma_1, x : V_c}{t_t}{U_t}{\Gamma_2, x : V_t}$ and that
$\typerule{\Gamma_1, x : V_c}{t_f}{U_f}{\Gamma_3, x : V_f}$ with
$\Gamma' = \Gamma_2 \sqcup \Gamma_3$, $U = U_t \sqcup U_f$ and
$V = V_t \sqcup V_f$.

By induction, there exists a $V_c' <: \remapapp{T'}{T}{V_c}$
such that $\typerule{\Gamma, x : T'}{t_c}{\boolt}{\Gamma_1, x : V_c'}$.
By \lemref{lem:remap} we also have that $V_c' <: V_c$.
Therefore by induction there also exists a 
$V_t' <: \remapapp{V_c'}{V_c}{V_t}$ and $V_f' <: \remapapp{V_c'}{V_c}{V_f}$
such that $\typerule{\Gamma_1, x : V_c'}{t_t}{T_t}{\Gamma_2, x : V_t'}$
and that $\typerule{\Gamma_1, x : V_c'}{t_f}{T_f}{\Gamma_3, x : V_f'}$.
By \lemref{lem:remap} we also have that $V_t' <: V_t$ and that
$V_f' <: V_f$.
Let $V' = V_t' \sqcup V_f'$. By the definition of join, $V' <: V$.
It follows by rule \tif that
$\typerule{\Gamma, x : T'}{t}{U}{\Gamma', x : V'}$.

\end{itemize}
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% lem:weakening %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{lem}[Weakening]
\label{lem:weakening}
If $\typerule{\Gamma}{t}{T}{\Gamma'}$,
then for all $\Gamma''$ such that $dom(\Gamma) \cap dom(\Gamma'') = \emptyset$,
$t$ can also be typed such that
$\typerule{\Gamma, \Gamma''}{t}{T}{\Gamma', \Gamma''}$.
\end{lem}
\begin{proof}
by induction on the structure of $t$.

If $t$ is some value $v$, it follows by \lemref{lem:valuectx} that 
$\Gamma = \Gamma'$ and that the typing derivation will still be valid as long
as the input and output contexts are identical. Therefore,
$\typerule{\Gamma , \Gamma''}{v}{T}{\Gamma', \Gamma''}$.

If $t$ is not a value, then it is of one of the following forms:

\begin{itemize}
\item $t = \lett{x}{t'}{t''}$. It follows from rule \tlet that
there exists $\Gamma_1, \Gamma_2, T', T''$ such that
$\typerule{\Gamma}{t'}{T'}{\Gamma_1}$ and
$\typerule{\Gamma_1, x : T'}{t''}{T}{\Gamma', x : T''}$.
By \tlet we have that $dom(\Gamma') = dom(\Gamma)$.
By induction, $t'$ can be typed such
that $\typerule{\Gamma,\Gamma''}{t'}{T'}{\Gamma_1,\Gamma''}$,
and $t''$ can be typed such that
$\typerule{\Gamma_1, \Gamma'', x : T'}{t''}{T}{\Gamma', \Gamma'', x : T''}$,
as we can assume by Barendregt's convention that $x$ can be made distinct from
any variable names in $\Gamma''$ by relabeling $x$.
Therefore,
$\typerule{\Gamma, \Gamma''}{\lett{x}{t'}{t''}}{T}{\Gamma', \Gamma''}$.

\item $t = x ( \overline{x_i} )$. It follows from the rule \tfunc that
there exists a $\Gamma_1, \overline{T_i}, \overline{U_i}, \overline{V_i}, \overline{T_i'}$ 
such that 
$\Gamma = \Gamma_1, x : (\overline{U_i \gg V_i}) \rightarrow T, \overline{x_i : T_i}$
and that
$\Gamma' = \Gamma_1, x : (\overline{U_i \gg V_i}) \rightarrow T, \overline{x_i : T_i'}$.
As $\Gamma_1$ can be arbitrary in the rule \tfunc, it follows that we
can extend the input and output contexts such that
$\typerule{\Gamma, \Gamma''}{x ( \overline{x_i} )}{T}{\Gamma', \Gamma''}$.

\item $t = x.m$. It follows from the rule \tmethc that there exists
a $\Gamma_1$ such that $\Gamma = \Gamma_1, x : O@\overline{S}$ and that 
$\Gamma' = \Gamma_1, x : O@\overline{S'}$.
As $\Gamma_1$ can be arbitrary in rule \tmethc, it follows that we can
extend the input and out contexts such that
$\typerule{\Gamma, \Gamma''}{x.m}{T}{\Gamma', \Gamma''}$.

\item $t = t' ; t''$. It follows from the rule \tseq that there exists
$\Gamma_1$ and $T'$ such that $\typerule{\Gamma}{t'}{T'}{\Gamma_1}$
and $\typerule{\Gamma_1}{t''}{T}{\Gamma'}$. By induction
$\typerule{\Gamma, \Gamma''}{t'}{T'}{\Gamma_1, \Gamma''}$
as $dom(\Gamma_1) \subseteq dom(\Gamma)$ and therefore
$\Gamma_1$ and $\Gamma''$ are disjoint.
Similarly, as $dom(\Gamma_1) = dom(\Gamma')$, by induction
$\typerule{\Gamma_1, \Gamma''}{t''}{T}{\Gamma', \Gamma''}$.
Therefore $\typerule{\Gamma, \Gamma''}{t'}{T'}{\Gamma_1, \Gamma''}$.

\item $t = \ift{t_c}{t_t}{t_f}$. It follows from rule \tif that there
exists $\Gamma_1$, $\Gamma_2$, $\Gamma_3$, $T_t$ and $T_f$ such that
$\typerule{\Gamma}{t_c}{\boolt}{\Gamma_1}$ and that
$\typerule{\Gamma_1}{t_t}{T_t}{\Gamma_2}$ and that
$\typerule{\Gamma_1}{t_f}{T_f}{\Gamma_3}$ with
$\Gamma' = \Gamma_2 \sqcup \Gamma_3$ and $T = T_t \sqcup T_f$.
By induction each of $\Gamma_1$, $\Gamma_2$ and $\Gamma_3$ can
be extended by $\Gamma''$ in the typing judgements for $t_c$,
$t_t$ and $t_f$. As $\Gamma'' \sqcup \Gamma'' = \Gamma''$, it
follows that $(\Gamma_2, \Gamma'') \sqcup (\Gamma_3, \Gamma'') = 
\Gamma_2 \sqcup \Gamma_3, \Gamma''$ and therefore $t$ can be
typed such that $\typerule{\Gamma, \Gamma''}{t}{T}{\Gamma', \Gamma''}$.

\end{itemize}

\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% lem:subst %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{lem}[Substitution]
\label{lem:subst}
If $\typerule{\Gamma}{t}{T}{\Gamma'}$, then
$\typerule
{\subst{x_i}{y_i}{\Gamma}}
{\subst{x_i}{y_i}{t}}
{T}
{\subst{x_i}{y_i}{\Gamma'}}$
where each $x_i$ and $y_i$ is distinct, and $\overline{x_i} \cap \overline{y_i} = \emptyset$.
\end{lem}
\begin{proof}
by induction on the structure of $t$.

If $t$ is some value $v$, then $\Gamma = \Gamma'$, therefore
$\subst{x_i}{y_i}{\Gamma} = \subst{x_i}{y_i}{\Gamma'}$
and substitution has no effect on the value, which has no specific requirements
of either $\Gamma$ or $\Gamma'$ other than that they are equal.
Therefore, $\typerule{\subst{x_i}{y_i}{\Gamma}}{\subst{x_i}{y_i}{v}}{T}{\subst{x_i}{y_i}{\Gamma'}}$.

If $t$ is a term, then it is of one of the following forms:

\begin{itemize}
\item $t = \lett{x}{t''}{t'}$. 
Substitution is defined on this term such that \\
$\subst{x_i}{y_i}{t} \equiv 
\lett{x}{( \subst{x_i}{y_i}{t''} )}{(\subst{x_j}{y_j}{t'})}$
, where $\overline{y_j} = \overline{y_i} - \{ x \}$.

By rule \tlet, it follows
that there exists $T', T'', \Gamma_1$ such that
$\typerule{\Gamma}{t''}{T''}{\Gamma_1}$ and that
$\typerule{\Gamma_1, x : T''}{t''}{T}{\Gamma', x : T'}$.
Additionally, $x \notin dom(\Gamma)$.

By induction, $t''$ can be substituted such that 
$\typerule
{\subst{x_i}{y_i}{\Gamma}}
{\subst{x_i}{y_i}{t''}}
{\subst{x_i}{y_i}{\Gamma_1}}
$.
Additionally, $t'$ can be substituted such that
$(\subst{x_j}{y_j}{\Gamma_1}) , x : T'' \triangleright \subst{x_j}{y_j}{t''} \triangleleft \{ (\overline{x_j / y_j} \} \Gamma') , x : T'$ as it is guaranteed that $x \notin \overline{y_j}$.
As $dom(\Gamma') = dom(\Gamma)$, and that $x \notin dom(\Gamma)$, it follows that $\subst{x_i}{y_i}{\Gamma_1} = (\subst{x_j}{y_j}{\Gamma_1})$ and
that $\overline{x_i / y_i} \} \Gamma' = \overline{x_j / y_j} \} \Gamma'$.

Therefore, the requirements of \tlet are satisfied such that
$\typerule{\subst{x_i}{y_i}{\Gamma}}{\subst{x_i}{y_i}{t}}{T}{\subst{x_i}{y_i}{\Gamma'}}$.

\item $t = t_l ; t_r$.
Substitution is defined on this term such that
$\{ \overline{x_i / y_i} \} ( t_l ; t_r ) \equiv \subst{x_i}{y_i}{t_l} ; \subst{x_i}{y_i}{t_r}$.

By rule \tseq, there exists $\Gamma'', T'$ such that
$\typerule{\Gamma}{t_l}{T'}{\Gamma''}$ and
$\typerule{\Gamma''}{t_r}{T}{\Gamma'}$. By induction,
$\typerule{\subst{x_i}{y_i}{\Gamma}}{\subst{x_i}{y_i}{t_l}}{T'}{\subst{x_i}{y_i}{\Gamma''}}$ and
$\typerule{\subst{x_i}{y_i}{\Gamma''}}{\subst{x_i}{y_i}{t_r}}{T}{\subst{x_i}{y_i}{\Gamma'}}$.
Therefore, $t$ can by typed \tseq after the substitution.

\item $t = x_0 ( \overline{x_k} )$.
Substition is defined on this term such that
$\subst{x_i}{y_i}{x_0} ( \overline{x_k} ) \equiv y_0 ( \overline{y_k} )$
where for each $x \in \{ x_0, \overline{x_k} \}$,
$x = y_j \implies y = x_j$, otherwise $y$ is the original value.

For each substitution that occurs in the term, the variable must exist in
both the input and output contexts by rule \tfunc, and will be substituted
for the same name. Therefore, $t$ can be typed by \tfunc after the
substitution.

\item $t = x.m$.
Substition is defined on this term such that
$\subst{x_i}{y_i}{x.m} \equiv y.m$
where $y = x_j$ if $x = y_j$ for some $y_j \in \overline{y_i}$.
Otherwise, $y = x$.

Rule \tmethc requires that $x$ exist in both $\Gamma$ and $\Gamma'$ and
therefore any substitution of x in the term will be matched with the
same substitution in the contexts. Therefore, $t$ can be typed by
\tmethc after the substitution.

\item $t = \ift{t_c}{t_t}{t_f}$. Substitution is defined on this term such
that \\
$\subst{x_i}{y_i}{(\ift{t_c}{t_t}{t_f})} \equiv 
\ift{\substxy{t_c}}{\substxy{t_t}}{\substxy{t_f}}$.
By rule \tif there exists $\Gamma_1$, $\Gamma_2$, $\Gamma_3$, $T_t$, $T_f$
such that
$\typerule{\Gamma}{t_c}{\boolt}{\Gamma_1}$ and that
$\typerule{\Gamma_1}{t_t}{T_t}{\Gamma_2}$ and that
$\typerule{\Gamma_1}{t_f}{T_f}{\Gamma_3}$ where
$\Gamma' = \Gamma_2 \sqcup \Gamma_3$ and $T = T_t \sqcup T_f$. By induction
substitution can be performed on $t_c$, $t_t$ and $t_f$ such that
$\typerule{\substxy{\Gamma}}{\substxy{t_c}}{\boolt}{\substxy{\Gamma_1}}$ and that
$\typerule{\substxy{\Gamma_1}}{\substxy{t_t}}{T_t}{\substxy{\Gamma_2}}$ and
$\typerule{\substxy{\Gamma_1}}{\substxy{t_f}}{T_f}{\substxy{\Gamma_3}}$.
By the definition of substitution on contexts, 
$\substxy{\Gamma_2} \sqcup \substxy{\Gamma_3} = \substxy{\Gamma'}$. Therefore,
by rule \tif, 
$\typerule{\substxy{\Gamma}}{\substxy{\ift{t_c}{t_t}{t_f}}}{T}{\substxy{\Gamma'}}$.
\end{itemize}
\end{proof}

\bibliographystyle{abbrvnat}
\bibliography{refs}

% The bibliography should be embedded for final submission.

%\begin{thebibliography}{}
%\softraggedright

%\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
%P. Q. Smith, and X. Y. Jones. ...reference text...

%\end{thebibliography}

\end{document}